---
title: "Dimensionality"
author: "Spencer Gray"
date: "2022-10-08"
output:
  html_document: default
  pdf_document: default
---

## Data Setup

Cleaning up our data, removing all NAs and setting to 0 to assist with our kNN model in the future.

```{r setup}
library(caret)

aus <- read.csv("weatherAUS.csv")

aus <- subset(aus,RainTomorrow  != "NA")

for(i in c(3, 4, 5, 9, 12, 13, 14, 15, 16, 17, 20, 21))
{
  aus[is.na(aus[,i]), i] <- mean(aus[,i], na.rm = TRUE)
}


dim(aus)
head(aus)

i <- sample(1:nrow(aus), 0.8 * nrow(aus), replace = FALSE)

train <- aus[i,]
test <- aus[-i,]
```

## Data Representation

Selecting relatively stable and numerically recorded variables (quantitative) to use our PCA model on. Predicting rain tomorrow in Column 23.
MinTemp       -> column 3
MaxTemp       -> column 4
Rainfall      -> column 5
WindGustSpeed -> column 9
WinSpeed9am   -> column 12
WinSpeed3pm   -> column 13
Humidity9am   -> column 14
Humidity3pm   -> column 15
Pressure9am   -> column 16
Pressure3pm   -> column 17
Temp9am       -> column 20
Temp3pm       -> column 21

```{r}
set.seed(1234)
pcaModel <- preProcess(train[,c(3, 4, 5, 9, 12, 13, 14, 15, 16, 17, 20, 21, 23)], method = c("center", "scale", "pca"))
pcaModel
```

## PCA Model Setup

```{r}
trainPCA <- predict(pcaModel, train[, c(3, 4, 5, 9, 12, 13, 14, 15, 16, 17, 20, 21, 23)])
testPCA <- predict(pcaModel,  test[, c(3, 4, 5, 9, 12, 13, 14, 15, 16, 17, 20, 21, 23)])
```

## PCA Model Accuracy

Graphing data of different principal components against one another. Training a knn model with PCA predict rain tomorrow based on variants of specific weather characteristics including: temperature, humidity, and wind.

```{r}
library(class)
set.seed(1234)

trainDF <- data.frame(trainPCA$PC1, trainPCA$PC2, trainPCA$PC3, trainPCA$PC4, trainPCA$PC5, trainPCA$PC6, trainPCA$PC7, train$RainTomorrow)
testDF <- data.frame(testPCA$PC1, testPCA$PC2, testPCA$PC3, testPCA$PC4, testPCA$PC5, testPCA$PC6, testPCA$PC7, test$RainTomorrow)



par(mfrow=c(3,2))
plot(trainPCA$PC1, trainPCA$PC2)
plot(trainPCA$PC2, trainPCA$PC3)
plot(trainPCA$PC3, trainPCA$PC4)
plot(trainPCA$PC4, trainPCA$PC5)
plot(trainPCA$PC5, trainPCA$PC6)
plot(trainPCA$PC6, trainPCA$PC7)

start_time <- Sys.time()
pred_reduced <- knn(trainDF[,1:7], testDF[,1:7], trainDF[,8], k = 6)

mean(pred_reduced == test$RainTomorrow)
```

## Regular Model Accuracy

Training a kNN model based on all the previous parameters but unmodified.

```{r}
library(class)
set.seed(1234)

trainDF_real <- data.frame(train$MinTemp, train$MaxTemp, train$WindGustSpeed, train$WindSpeed9am, train$WindSpeed3pm, train$Humidity9am, train$Humidity3pm, train$Pressure9am, train$Pressure3pm, train$Temp9am, train$Temp3pm, train$Rainfall, train$RainTomorrow)

testDF_real <- data.frame(test$MinTemp, test$MaxTemp, test$WindGustSpeed, test$WindSpeed9am, test$WindSpeed3pm, test$Humidity9am, test$Humidity3pm, test$Pressure9am, test$Pressure3pm, test$Temp9am, test$Temp3pm, test$Rainfall, test$RainTomorrow)


pred <- knn(trainDF_real[,1:12], testDF_real[,1:12], trainDF_real[,13], k = 10)

mean(pred==test$RainTomorrow)
```

## LDA Model Setup and Accuracy

```{r}
library(MASS)
ldaModel <- lda(RainTomorrow~MinTemp + MaxTemp + WindGustSpeed + WindSpeed9am + WindSpeed3pm + Humidity9am + Humidity3pm + Pressure9am + Pressure3pm + Temp9am + Temp3pm + Rainfall, data = train)
ldaModel$means


lda_pred <- predict(ldaModel, newdata=test, type="class")
mean(lda_pred$class==test$RainTomorrow)
plot(lda_pred$x[,1], lda_pred$posterior[,1])
```
